Profile likelihoods for geostatsistical models using GPU's

### 2022-01-05 pkg/oldgpuRandom/inst/doc/L2021_11/paper2.Rnw
paper 2
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3097 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

<<packages, results='hide', message=FALSE>>=
 library(gpuRandom)
 library(gpuR)
 library(data.table)
 #library(geostatsp)
@

@@

\section{A simulation study}

>>>>

<<packages, results='hide', message=FALSE>>=
 library(gpuRandom)
 library(gpuR)
 library(data.table)
 library(geostatsp)
@

@@

\section{A simulation study}
<<simulate_datasets, echo=FALSE, cache=TRUE, results='hide', message=FALSE>>=
############### A SIMULATION STUDY IN THE PAPER 2 ####################################
Nsim=25
set.seed(88)
mydat = SpatialPointsDataFrame(cbind(seq(100*1e3,140*1e3, len=500), seq(100*1e3,120*1e3, len=500)), 
                               data=data.frame(cov1 =stats::rnorm(500, 120, 20), cov2 = stats::rpois(500, 50))
)

## simulate a random field
trueParam = c(range=1000, shape=2, variance=0.5^2, tauSq=1^2, anisoRatio=1, anisoAngleRadians=0)
nugget = trueParam['tauSq']/trueParam['variance']
names(nugget) <- "nugget"
nugget   #4
boxcox = c(boxcox=1)
options("useRandomFields" = FALSE)
for (i in 1:Nsim){
  set.seed(i*6-1)
  U <- geostatsp::RFsimulate(model=trueParam[setdiff(names(trueParam),'tauSq')],x=mydat)@data
  # simulate response
  Y <- 3 + 1*mydat$cov1 + 0.5*mydat$cov2 + U + stats::rnorm(length(mydat), 0, sd=sqrt(trueParam["tauSq"]))
  mydat@data <- cbind(mydat@data,Y)
}
colnames(mydat@data) <- c("cov1", "cov2", paste("Y", c(1:Nsim), sep=""))

## geostatsp's estimates
swissRes =  lgm( formula=Y10~ cov1 + cov2,
                 data=mydat,
                 grid=20,
                 reml = FALSE,
                 boxcox=1,
                 fixBoxcox=TRUE, fixShape=FALSE, fixNugget = FALSE,  #Set to FALSE to estimate the nugget effect parameter.
                 aniso=FALSE )
swissRes$summary[,c('estimate','ci0.005', 'ci0.995')]
swissRes$optim$mle
@

<<set params, echo=FALSE, cache=TRUE>>=
newParamList = list(
  range = c(exp(seq(log(0.3), log(1), len=10))*1000, exp(seq(log(1.2), log(2.9), len=12))*1000),  
  nugget = c(seq(0, 4, len=8), seq(4.5, 20, len=12)),  
  shape = 2,
  anisoRatio =1,
  anisoAngleRadians = 0.0
) 
params = do.call(expand.grid, newParamList)
@

<<run_forloop, echo=FALSE, results='hide', message=FALSE, warning=FALSE, cache=TRUE>>=
count = 0
breaksf = rep(0, 10)
makedata <- rep(0, 4)
estimates <- rep(0, 6)

library(gpuRandom)
library(gpuR)
count = 0
breaksf = rep(0, 10)
makedata <- rep(0, 4)
estimates <- rep(0, 6)
## calculate logl
for (i in 1:Nsim){
  result<-gpuRandom::likfitLgmCov2d(
    data= mydat,
    formula= as.formula(paste(colnames(mydat@data)[i+2], "~", "cov1 + cov2")), 
    coordinates=mydat@coords,
    params=params, 
    paramToEstimate = c('range','nugget'),
    boxcox = 1,
    cilevel=0.8,
    type = "double",
    reml=FALSE, 
    NparamPerIter=400,
    Nglobal=c(128,64),
    Nlocal=c(16,16),
    NlocalCache=2800,
    verbose=FALSE)

  a <- geostatsp::loglikLgm(
    c(trueParam['range'],
      trueParam['shape'],
      nugget,
      trueParam[c('anisoRatio')],
      trueParam[c('anisoAngleRadians')],
      boxcox),
    data = mydat,
    formula = as.formula(paste(colnames(mydat@data)[i+2], "~", "cov1 + cov2")),
    reml = FALSE,
    minustwotimes=FALSE)[['logLik']]

  if (a  >= result$breaks){
    count = count +1
  }
  breaksf[i] = result$breaks
  makedata0 <- cbind(params[,1:2], result$LogLik, D=i)
  makedata <- rbind(makedata, makedata0)
  estimates <- cbind(estimates, result$summary)
}
colnames(makedata) <- c("range",'nugget', 'LogLik', 'D')
makedatahey <- makedata[-1,]
estimates_summary <- estimates[,-1]
mean_estimates <- apply(estimates_summary, 1, mean)
@

<<results, cache=TRUE>>=
count/Nsim
mean_estimates
@

<<graphs1, echo = FALSE, eval=TRUE, cache=TRUE>>=
library(RColorBrewer)
library('ggplot2')
forigin <- ggplot(subset(makedatahey, D==1), aes(x = range, y = nugget, z=LogLik)) + theme_bw() + 
  scale_y_continuous(limits=c(0,20))+
  scale_x_continuous(limits=c(300,3000) )+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  geom_point(size = 0.5, colour='grey') + 
  stat_contour(breaks=breaksf[1])  + 
  geom_vline(xintercept = trueParam['range'])+
  geom_hline(yintercept=nugget)
cols <- rainbow(Nsim/2)
for (i in 2:Nsim/2){
  forigin <- forigin + geom_contour(data = subset(makedatahey, D==i), aes(x = range, y = nugget, z=LogLik), breaks=breaksf[i], col= cols[i])
}
@
<<showgraphs1, echo = FALSE, eval=TRUE, cache=TRUE, fig.height =3, fig.width =4, fig.show ="hold",  fig.cap = "title", dev='png'>>=
forigin
@
